---
title: "Some Analysis on MTCARS dataSet"
output: pdf_document
---

#Executive summary

In this document, we are going to study and to analyze the possible relation between the transmission type of a car and its miles per gallon. In our study, we will perform the following steps :\newline
1. loading data and getting a preview\newline
2. exploring data\newline
3. fitting a regressive model\newline
4. checking if our model is robust when adding new variables.\newline
5. trying to conclude based on the results of our tests.\newline

#Preview

The data set we are going to study is "mtcars".

```{r} 
data("mtcars")
head(mtcars,2)
```

#Exploratory Analysis
let's divide and compare data into two groups of cars by the type of transmission: "auto" and "manuel" transmission.

```{r, echo=FALSE ,fig.width=12}
part1<-subset(mtcars,am==0)
part2<-subset(mtcars,am==1)
library(ggplot2)
library(gridExtra)
library(grid)
library(lattice)

x1<-ggplot(aes(y = mpg, x = factor(am),fill=factor(am)), data = mtcars) + geom_boxplot()+ theme_bw()
x2<-ggplot(aes(y = cyl, x = factor(am),fill=factor(am)), data = mtcars) + geom_boxplot()+ theme_bw()
x3<-ggplot(aes(y = disp, x = factor(am),fill=factor(am)), data = mtcars) + geom_boxplot()+ theme_bw()
x4<-ggplot(aes(y = hp, x = factor(am),fill=factor(am)), data = mtcars) + geom_boxplot()+ theme_bw()
x5<-ggplot(aes(y = drat, x = factor(am),fill=factor(am)), data = mtcars) + geom_boxplot()+ theme_bw()
x6<-ggplot(aes(y = wt, x = factor(am),fill=factor(am)), data = mtcars) + geom_boxplot()+ theme_bw()
x7<-ggplot(aes(y = qsec, x = factor(am),fill=factor(am)), data = mtcars) + geom_boxplot()+ theme_bw()
x8<-ggplot(aes(y = gear, x = factor(am),fill=factor(am)), data = mtcars) + geom_boxplot()+ theme_bw()
x9<-ggplot(aes(y = carb, x = factor(am),fill=factor(am)), data = mtcars) + geom_boxplot()+ theme_bw()

grid.arrange(x1, x2, x3, x4,x5,x6,x7,x8,x9, ncol=5)

```

We can see that mpg are differents in the subgroups of transmission mode.
However,we can see also that generally the two groups don't have the same characteristics : This means that we have to be careful before stating that only type transmission affects the mpg -> there might other factors that cause the mpg changes.

#Inference  

Let's check if subgrouping by type of transmission creates samples with different means :

```{r, echo=TRUE ,fig.width=10}
ks.test(mtcars[mtcars$am==1,]$mpg,mtcars[mtcars$am==0,]$mpg,"pgamma")
```

with a p-value of 0.003911 , we are pretty sure that the two subgroups have different means and we can infer it. 

# Fitting a model : mpg Vs am
We know that mpg is positive and isn't always an integer. Naturally, we can't think of using the Gamma model.
Lets compare the residuals of a normal(classical)linear model with a generalised linear model using a gamma family.

```{r, echo=FALSE ,fig.width=10, fig.height=2}
mdlN<-lm(mpg~factor(am)-1,data=mtcars)
mdlG<-glm(mpg~factor(am)-1,data=mtcars,family=Gamma)

residG<-resid(mdlG)
residN<-resid(mdlN)
indexI<-c(rep("Normal Model",length(resid((mdlG)))),rep("Gamma Model",length(resid((mdlG)))))

residValues<-c(residN,residG)

df<-data.frame(residValues,indexI)
ggplot(aes(y = residValues, x = factor(indexI),fill=factor(indexI)), data = df) + geom_boxplot()+ theme_bw()
```

Gamma model seems to fit greatly the data. 

### What if .. ? other variables might explain the difference 

Let's take a look at the coefficients of our gamma model :

```{r, echo=FALSE ,fig.width=10}
mdlG$coefficients
```

Let's see what happens if we add the variable weight :

```{r, echo=FALSE ,fig.width=10}
mdlG2<-glm(mpg~factor(am)+wt -1,data=mtcars,family=Gamma)
mdlG2$coefficients
```

now let's add weight and cylinders too :
```{r, echo=FALSE ,fig.width=10}
mdlG3<-glm(mpg~factor(am)+wt + cyl -1,data=mtcars,family=Gamma)
mdlG3$coefficients

```
Yes , we can see that including "cyl" and "wt"" has changed radically the coefficients of the factors of transmission mod.

Let's compare the residuals of the three models

```{r, echo=FALSE ,fig.width=10}
residG1<-resid(mdlG)
residG2<-resid(mdlG2)
residG3<-resid(mdlG3)
numberParams<-c(rep("Gamma 1 param",length(resid((mdlG)))),
           rep("Gamma 2 params",length(resid((mdlG2)))),
           rep("Gamma 3 params",length(resid((mdlG3)))))

residualValues<-c(residG1,residG2,residG3)

df1<-data.frame(residualValues,numberParams)
ggplot(aes(y = residualValues, x = factor(numberParams),fill=factor(numberParams)), data = df1) + geom_boxplot()+ theme_bw()

sums<-c(sum(residG1^2),sum(residG2^2),sum(residG3^2))
sums
```

as we can see, introducing the new variables "cyl" and "wt" has reduced the residuals and improved the quality of our model.

let's compare the three models using anova :
```{r, echo=FALSE ,fig.width=10}
anova(mdlG,mdlG2,mdlG3,test="Chisq")
```

Anova confirms that the additional variables bring information.


# Conclusion
The analysis above show that we have no evidence that the transmission mode is responsable of mpg value. To make a more valuable analysis , It might be better to have data that compares cars that differ only in mode of transmission and are similar in other characterisics(than mpg and am).


#Appendix
1) To produce the exploratory figures ,  i used the following code:
```{r ,fig.width=12, fig.lp="hello world"}
part1<-subset(mtcars,am==0)
part2<-subset(mtcars,am==1)
# please uncomment the four following lines
#library(ggplot2)
#library(gridExtra)
#library(grid)
#library(lattice)

x1<-ggplot(aes(y = mpg, x = factor(am),fill=factor(am)), data = mtcars) + geom_boxplot()+ theme_bw()
x2<-ggplot(aes(y = cyl, x = factor(am),fill=factor(am)), data = mtcars) + geom_boxplot()+ theme_bw()
x3<-ggplot(aes(y = disp, x = factor(am),fill=factor(am)), data = mtcars) + geom_boxplot()+ theme_bw()
x4<-ggplot(aes(y = hp, x = factor(am),fill=factor(am)), data = mtcars) + geom_boxplot()+ theme_bw()
x5<-ggplot(aes(y = drat, x = factor(am),fill=factor(am)), data = mtcars) + geom_boxplot()+ theme_bw()
x6<-ggplot(aes(y = wt, x = factor(am),fill=factor(am)), data = mtcars) + geom_boxplot()+ theme_bw()
x7<-ggplot(aes(y = qsec, x = factor(am),fill=factor(am)), data = mtcars) + geom_boxplot()+ theme_bw()
x8<-ggplot(aes(y = gear, x = factor(am),fill=factor(am)), data = mtcars) + geom_boxplot()+ theme_bw()
x9<-ggplot(aes(y = carb, x = factor(am),fill=factor(am)), data = mtcars) + geom_boxplot()+ theme_bw()
 # please , uncomment the following line :
#grid.arrange(x1, x2, x3, x4,x5,x6,x7,x8,x9, ncol=5)

```
2)To fit regressive models , i used the following code :
```{r, fig.width=10, fig.height=2}
mdlN<-lm(mpg~factor(am)-1,data=mtcars)
mdlG<-glm(mpg~factor(am)-1,data=mtcars,family=Gamma)

residG<-resid(mdlG)
residN<-resid(mdlN)
indexI<-c(rep("Normal Model",length(resid((mdlG)))),rep("Gamma Model",length(resid((mdlG)))))

residValues<-c(residN,residG)

df<-data.frame(residValues,indexI)
# please , uncomment the following lines
#ggplot(aes(y = residValues, x = factor(indexI),fill=factor(indexI)), data = df) + geom_boxplot()+ theme_bw()
```
3) To add new variables into model and use anova , i used the following code :
```{r, fig.width=10, fig.height=2}

mdlG2<-glm(mpg~factor(am)+wt -1,data=mtcars,family=Gamma)
mdlG3<-glm(mpg~factor(am)+wt + cyl -1,data=mtcars,family=Gamma)
# please ,  uncomment the following lines
#mdlG$coefficients
#mdlG2$coefficients
#mdlG3$coefficients
#anova(mdlG,mdlG2,mdlG3,test="Chisq")
```